---
title: "Homework #3"
author: "Zoe M."
date: "2023-02-01"
output: 
  html_document: 
    highlight: textmate
    theme: cerulean
---

### Problem #1:

First String    Second      1.22      3.4

Second          More Text   1.555555  2.2220

Third           x           3         124

```{r,eval=FALSE}
Find: (\s{2,})
Replace: ,
```

First String,Second,1.22,3.4

Second,More Text,1.555555,2.2220

Third,x,3,124

### Problem #2:

Ballif, Bryan, University of Vermont

Ellison, Aaron, Harvard Forest

Record, Sydne, Bryn Mawr

```{r,eval=FALSE}
Find: (\w+),\s(\w+),\s(.*)
Replace: \1 \2 (\3)
```
Ballif Bryan (University of Vermont)

Ellison Aaron (Harvard Forest)

Record Sydne (Bryn Mawr)

### Problem #3:

0001 Georgia Horseshoe.mp3 0002 Billy In The Lowground.mp3 0003 Winder Slide.mp3 0004 Walking Cane.mp3

```{r,eval=FALSE}
Find: (\.\w+\d)\s
Replace: \1\n
```

0001 Georgia Horseshoe.mp3

0002 Billy In The Lowground.mp3

0003 Winder Slide.mp3

0004 Walking Cane.mp3

### Problem #4:

0001 Georgia Horseshoe.mp3

0002 Billy In The Lowground.mp3

0003 Winder Slide.mp3

0004 Walking Cane.mp3

```{r,eval=FALSE}
Find: (\d{4})\s(.*)(\.\w+\d)
Replace:\2_\1\3
```

Georgia Horseshoe_0001.mp3

Billy In The Lowground_0002.mp3

Winder Slide_0003.mp3

Walking Cane_0004.mp3

### Problem #5:

Camponotus,pennsylvanicus,10.2,44

Camponotus,herculeanus,10.5,3

Myrmica,punctiventris,12.2,4

Lasius,neoniger,3.3,55

```{r,eval=FALSE}
Find: (\w)\w+,(\w+),\d+.\d,(\d+)
Replace:\1_\2,\3
```

C_pennsylvanicus,44

C_herculeanus,3

M_punctiventris,4

L_neoniger,55

### Problem #6:

Camponotus,pennsylvanicus,10.2,44

Camponotus,herculeanus,10.5,3

Myrmica,punctiventris,12.2,4

Lasius,neoniger,3.3,55

```{r,eval=FALSE}
Find: (\w)\w+,(\w{4})\w+,\d+.\d,(\d+)
Replace:\1_\2,\3
```

C_penn,44

C_herc,3

M_punc,4

L_neon,55

### Problem #7:

Camponotus,pennsylvanicus,10.2,44

Camponotus,herculeanus,10.5,3

Myrmica,punctiventris,12.2,4

Lasius,neoniger,3.3,55

```{r,eval=FALSE}
Find: (\w{3})\w+,(\w{3})\w+,(\d+.\d),(\d+)
Replace:\1\2, \4, \3
```

Campen, 44, 10.2

Camher, 3, 10.5

Myrpun, 4, 12.2

Lasneo, 55, 3.3


