---
title: "Homework_6"
author: "Zoe M."
date: "2023-03-01"
output: 
  html_document: 
    highlight: textmate
    theme: cerulean
---

``` {r, echo=TRUE, message=FALSE}
# Library Calls:
library(tidyverse)
library(dplyr)
```

## Problem #1:

```{r, echo=TRUE}
data("iris")
class(iris)
```
The data set iris has 150 observations of 5 variables.

## Problem #2:

Create a new data frame iris1 that contains only the species virginica and versicolor with sepal lengths longer than 6 cm and sepal widths longer than 2.5 cm. How many observations and variables are in the data set?

```{r,echo=TRUE}
# using filter() to create data frame with specified conditions
iris1<-filter(iris, Species == c("virginica","versicolor") & Sepal.Length>6 & Sepal.Width>2.5)
print(iris1)
```
There are 28 observations of 5 variables in iris1.

## Problem #3:

Now, create a iris2 data frame from iris1 that contains only the columns for Species, Sepal.Length, and Sepal.Width. How many observations and variables are in the data set?

```{r, echo=TRUE}
# using select() to create data frame with 3 chosen columns
iris2<-select(iris1, Species, Sepal.Length, Sepal.Width) 
print(iris2)
```
There are 28 observations of 3 variables in iris2.

## Problem #4:

Create an iris3 data frame from iris2 that orders the observations from largest to smallest sepal length. Show the first 6 rows of this data set.

```{r, echo=TRUE}
# using arrange() to order observations of Sepal.Length by descending value
iris3<-arrange(iris2, by=desc(Sepal.Length))
# show first 6 rows of iris3 
head(iris3) 
```

## Problem #5:

Create an iris4 data frame from iris3 that creates a column with a sepal area (length * width) value for each observation. How many observations and variables are in the data set?

```{r, echo=TRUE}
# using mutate() to add a new column, fill column with calculated sepal areas for each observation
iris4<-mutate(iris3, Sepal.Area=(Sepal.Length*Sepal.Width))
print(iris4)
```

There are 28 observations of 4 variables in iris4.

## Problem #6:

Create iris5 that calculates the average sepal length, the average sepal width, and the sample size of the entire iris4 data frame and print iris5.

```{r, echo=TRUE}
# using summarize() to create data frame with mean sepal length, mean sepal width, and sample size of data frame
iris5<-summarize(iris4, 
                 Mean.Sepal.Length=mean(Sepal.Length), 
                 Mean.Sepal.Width=mean(Sepal.Width),
                 number=n())
print(iris5)
```

## Problem #7:

Finally, create iris6 that calculates the average sepal length, the average sepal width, and the sample size for each species of in the iris4 data frame and print iris6.

```{r, echo=TRUE}
# same use of summarize() as above, but using .by=Species to group summary by species
iris6<-summarize(iris4, 
            Mean.Sepal.Length=mean(Sepal.Length), 
            Mean.Sepal.Width=mean(Sepal.Width), 
            .by=Species, 
            number=n())
print(iris6)
```


## Problem #8:

See if you can rework all of your previous statements (except for iris5) into an extended piping operation that uses iris as the input and generates irisFinal as the output.

```{r, echo=TRUE}
irisFinal<-iris%>%
  filter(Species == c("virginica","versicolor") & Sepal.Length>6 & Sepal.Width>2.5)%>%
  select(Species, Sepal.Length, Sepal.Width)%>%
  arrange(by=desc(Sepal.Length))%>%
  mutate(Sepal.Area=Sepal.Length/Sepal.Width)%>%
  summarize(Mean.Sepal.Length=mean(Sepal.Length), 
            Mean.Sepal.Width=mean(Sepal.Width), 
            .by=Species, 
            number=n())
print(irisFinal)
```

## Problem #9:

Create a ‘longer’ data frame using the original iris data set with three columns named “Species”, “Measure”, “Value”. The column “Species” will retain the species names of the data set. The column “Measure” will include whether the value corresponds to Sepal.Length, Sepal.Width, Petal.Length, or Petal.Width and the column “Value” will include the numerical values of those measurements.

```{r, echo=TRUE}
# using pivot_longer to take column names from Sepal.Length:Petal.Width, make these column names values in a new column called "Measure", 
iris_longer<-iris%>%
  pivot_longer(cols=Sepal.Length:Petal.Width, names_to="Measure", values_to="Value", values_drop_na=T)
print(iris_longer)
```


