---
title: "Homework_12"
author: "Zoe M."
date: "2023-05-08"
output: 
  html_document: 
    highlight: textmate
    theme: cerulean
---

# Abstract:

The emerald ash borer (Agrilus planipennis) is a species of native to Northern Asia and highly invasive outside of its native regions. In North America, these beetles are highly invasive and cause mass destruction of ash tree forests. Tetrastichus planipennisi, a parasite of A. planipennis, has achieved success as a biological control against these infestations. Our research team at the University of Vermont sought to examine the changes in T. planipennisi morphological characteristics in covered and uncovered conditions while overwintering in the Jericho Research Forest in Vermont, USA. The wasp larvae (graciously received from the USDA-ARS BIIRU Laboratory) were reared in open air jars containing ash tree branches throughout 5 months of the winter of 2021 to 2022. Adult wasps were collected after the experiment and classified as being found inside or outside of the ash tree branches. Images of the wasps were taken using Leica Biosystems software and morphometric measurements were obtained using ImageJ software. The data were visualized using R studio version 4.2.1 to create histograms, box plots, and scatter plots. Analyses of variance found a statistically significant difference in the mean body (p=0.003) and notum (p=0.00216) lengths between covered and uncovered treatment groups, but did not find a significant difference in the mean right wing (p=0.0625) or right hind tibia (p=0.0818) lengths. Analyses of variance did not find a statistically significant difference in the mean right wing (0.0625), body (p=0.003), notum (p=0.00216), or right hind tibia (p=0.0818) lengths between wasps found inside and outside of the branches. Linear regression was used to assess the statistical relationships between body length and right wing (p<0.0001), notum (p<0.0001), and right hind tibia (p<0.0001) lengths. Each regression found a moderate, positive correlation between body length and the other major morphometrics. Our results help us gain a preliminary understanding of the morphological changes T. planipennisi experience in Vermont winters, and help inform further research into the use of the species as a biological control against A. planipennis infestations.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

# Initialization:
```{r}
library(ggplot2)
library(ggthemes)
library(patchwork)
library(hrbrthemes)
```

# File Preparation:
```{r}
# cleaned data with \d*\-\d\(3\)\w#\d*
m_treat <- read.csv("Morphometric_Covered_Uncovered.csv")
# cleaned data with \d*\-\d\(\d\)I#\d* or \d*\-\d\(\d\)O#\d*
m_found <- read.csv("Morphometric_Inside_Out.csv")
```

# Histograms:
```{r}
# histograms (overlap covered and uncovered)
h1 <- ggplot(data=m_treat) +
  aes(x = Right_Wing_Length, fill = X) +
  geom_histogram(position = "identity", 
                 alpha = 0.5, 
                 bins = 50, 
                 color="grey20") +
  xlab("Right Wing Length (mm)")+
  ylab("Count") +
  ggtitle("Right Wing Length Frequency")
# print(h1)

h2 <- ggplot(data=m_treat) +
  aes(x = Total_Body_Length, fill = X) +
  geom_histogram(position = "identity", 
                 alpha = 0.5, 
                 bins = 50,
                 color="grey20") +
  xlab("Total Body Length (mm)")+
  ylab("Count") +
  ggtitle("Total Body Length Frequency")
# print(h2)

h3 <- ggplot(data=m_treat) +
  aes(x = Notum_Length, fill = X) +
  geom_histogram(position = "identity", 
                 alpha = 0.5, 
                 bins = 50,
                 color="grey20") +
  xlab("Notum Length (mm)")+
  ylab("Count") +
  ggtitle("Notum Length Frequency")
# print(h3)

h4 <- ggplot(data=m_treat) +
  aes(x = Right_Hind_Tibia_Length, fill = X) +
  geom_histogram(position = "identity", 
                 alpha = 0.5, 
                 bins = 50,
                 color="grey20") +
  xlab("Right Hind Tibia Length (mm)")+
  ylab("Count") +
  ggtitle("Right Hind Tibia Length Frequency")
# print(h4)

(h1|h2)/(h3|h4)
```

# Box Plots (Covered vs. Uncovered):
```{r}
# boxplots (covered vs uncovered)
g1 <- ggplot(data=m_treat) +
  aes(x=X,
      y=Right_Wing_Length,) +
  xlab("Treatment Group") +
  ylab("Right Wing Length (mm)")+
  geom_boxplot(fill= c("#00BFC4", "#F8766D"), color="grey20") + 
  theme(legend.position="none")
# print(g1)

g2<-ggplot(data=m_treat) +
  aes(x=X,
      y=Total_Body_Length,) +
  xlab("Treatment Group") +
  ylab("Body Length (mm)")+
  geom_boxplot(fill= c("#00BFC4", "#F8766D"), color="grey20") + 
  theme(legend.position="none")
# print(g2)

g3<-ggplot(data=m_treat) +
  aes(x=X,
      y=Notum_Length,) +
  xlab("Treatment Group") +
  ylab("Notum Length (mm)")+
  geom_boxplot(fill= c("#00BFC4", "#F8766D"), color="grey20") + 
  theme(legend.position="none")
# print(g3)

g4<-ggplot(data=m_treat) +
  aes(x=X,
      y=Right_Hind_Tibia_Length,) +
  xlab("Treatment Group") +
  ylab("Right Hind Tibia Length (mm)")+
  geom_boxplot(fill= c("#00BFC4", "#F8766D"), color="grey20") + 
  theme(legend.position="none")
# print(g4)

(g1|g2)/(g3|g4)
```

# ANOVA (Covered vs. Uncovered)
```{r}
aov_rwing_treat <- aov(Right_Wing_Length ~ X, m_treat)
summary(aov_rwing_treat) # not significant 0.0625
aov_body_treat <- aov(Total_Body_Length ~ X, m_treat)
summary(aov_body_treat) # significant 0.003
aov_notum_treat <- aov(Notum_Length ~ X, m_treat)
summary(aov_notum_treat) # significant 0.00216
aov_tibia_treat <- aov(Right_Hind_Tibia_Length ~ X, m_treat)
summary(aov_tibia_treat) # not significant 0.0818
```

# Box Plots (Inside vs. Outside):
```{r}
b1 <- ggplot(data=m_found) +
  aes(x=ID,
      y=Right_Wing_Length,) +
  xlab("Treatment Group") +
  ylab("Right Wing Length (mm)")+
  geom_boxplot(fill= c("#C77CFF", "#7CAE00"), color="grey20") + 
  theme(legend.position="none")
# print(b1)

b2<-ggplot(data=m_found) +
  aes(x=ID,
      y=Total_Body_Length,) +
  xlab("Treatment Group") +
  ylab("Body Length (mm)")+
  geom_boxplot(fill= c("#C77CFF", "#7CAE00"), color="grey20") + 
  theme(legend.position="none")
# print(b2)

b3<-ggplot(data=m_found) +
  aes(x=ID,
      y=Notum_Length,) +
  xlab("Treatment Group") +
  ylab("Notum Length (mm)")+
  geom_boxplot(fill= c("#C77CFF", "#7CAE00"), color="grey20") + 
  theme(legend.position="none")
# print(b3)

b4<-ggplot(data=m_found) +
  aes(x=ID,
      y=Right_Hind_Tibia_Length,) +
  xlab("Treatment Group") +
  ylab("Right Hind Tibia Length (mm)")+
  geom_boxplot(fill= c("#C77CFF", "#7CAE00"), color="grey20") + 
  theme(legend.position="none")
# print(b4)

(b1|b2)/(b3|b4)
```

# ANOVA (Inside vs. Outside):
```{r}
aov_rwing_found <- aov(Right_Wing_Length ~ ID, m_found)
summary(aov_rwing_found) # not significant 0.901
aov_body_found <- aov(Total_Body_Length ~ ID, m_found)
summary(aov_body_found) # not significant 0.645
aov_notum_found <- aov(Notum_Length ~ ID, m_found)
summary(aov_notum_found) # not significant 0.84
aov_tibia_found <- aov(Right_Hind_Tibia_Length ~ ID, m_found)
summary(aov_tibia_found) # not significant 0.275
```

# Linear Regressions:
```{r}
m1 <- ggplot(data=m_treat) + 
  aes(x=Total_Body_Length,
      y=Right_Wing_Length,
      color=X) +
  ggtitle("Effect of Body Length on Right Wing Length")+ 
  xlab("Body Length (mm)") +
  ylab("Right Wing Length (mm)") +
  geom_point(size=2) +
  geom_smooth(method="lm",
              se=F,
              formula = y ~ x,
              color="grey20") +
  annotate("text", x=2.2, y=3.5, label="r = 0.6109") +
  annotate("text", x=2.2, y=3.3, label="r^2 = 0.3732") +
  annotate("text", x=2.2, y=3.1, label="p = 3.133e-06")
m1
# lm2 <- lm(Right_Wing_Length~Total_Body_Length, data=m_treat)
# summary(lm1)

m2 <- ggplot(data=m_treat) + 
  aes(x=Total_Body_Length,
      y=Notum_Length,
      color=X) +
  ggtitle("Effect of Body Length on Notum Length")+ 
  xlab("Body Length (mm)") +
  ylab("Notum Length (mm)") +
  geom_point(size=2) +
  geom_smooth(method="lm",
              se=F,
              formula = y ~ x,
              color="grey20") +
  annotate("text", x=2.2, y=1.5, label="r = 0.5560") +
  annotate("text", x=2.2, y=1.4, label="r^2 = 0.3091") +
  annotate("text", x=2.2, y=1.3, label="p = 7.342e-05")
m2
# lm2 <- lm(Notum_Length~Total_Body_Length, data=m_treat)
# summary(lm2)

m3 <- ggplot(data=m_treat) + 
  aes(x=Total_Body_Length,
      y=Right_Hind_Tibia_Length,
      color=X) +
  ggtitle("Effect of Body Length on Right Hind Tibia Length")+ 
  xlab("Body Length (mm)") +
  ylab("Right Hind Tibia Length (mm)") +
  geom_point(size=2) +
  geom_smooth(method="lm",
              se=F,
              formula = y ~ x,
              color="grey20") +
  annotate("text", x=2.2, y=1.3, label="r = 0.6692") +
  annotate("text", x=2.2, y=1.2, label="r^2 = 0.4478") +
  annotate("text", x=2.2, y=1.1, label="p = 3.173e-08")
m3
# lm3 <- lm(Right_Hind_Tibia_Length~Total_Body_Length, data=m_treat)
# summary(lm3)
```

